---
title: "Qualitative use-wear analysis"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```

---


# Goal of the script
This script formats the output of the resulting CSV-file from digitalising the location of use-wear traces in QGIS.
The script will:

1. Read in the original CSV-file   
2. Format the data   
3. Write an XLSX-file and save an R object ready for further analysis in R 

```{r}
dir_in <- "analysis/raw_data"
dir_out <- "analysis/derived_data/"
```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(tools)
library(R.utils)
library(chron)
library(ggplot2)
library(wesanderson)
library(dplyr)
library(doBy)
```


---

# Get names, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.csv$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
```

The checksum (MD5 hashes) of the imported file are:  
```{r, echo = FALSE}
info_in
```


---


# Read in original CSV-file
```{r}
imp_data <- read.csv(data_file, header = TRUE, stringsAsFactors = FALSE, na.strings = "*****")
str(imp_data)
```

---


# Pie charts
```{r}


# define which part of the data belongs to Keilmesser 
KM <- imp_data[imp_data[["tool.type"]] == "Keilmesser", ]

# splits the data into single areas 
sp <- split(KM, KM[["area"]])


custom.col <- c("#999999", "#FFDB6D", "#C4961A", "#9a0f0f", "#F4EDCA", "black",
                "#D16103", "#C3D7A4", "#52854C", "#4E84C4", "#293352") 

pct.fun <- function(x, y) length(x) / y *100
n.rows <- summaryBy(specimen ~ area, data = KM, FUN = length)
for (i in seq_along(n.rows)) {
  pct[[i]] <- summaryBy(specimen ~ area + use.wear.type, data = KM, FUN = pct.fun, y = n.rows[i])
}
pct <- summaryBy(specimen ~ area + use.wear.type, data = KM, FUN = pct.fun, y = KM)


# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp)){
  KM_pie <- ggplot(data = sp[[i]], aes(x = area, fill = use.wear.type )) +
            geom_bar(stat = "count", width = 0.5) + 
            coord_polar("y", start = 0) +
            theme_void() +
            #facet_wrap(~area) +
            scale_fill_manual(values = custom.col) +
            labs(fill = gsub("\\.", " ", "use-wear type"))  
            #geom_text(aes(y = use.wear.type), color = "white")
          
 
  print(KM_pie)

}

#F4EDCA



```

