---
title: "Qualitative use-wear analysis"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output: word_document
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)

```

---


# Goal of the script
This script formats the output of the resulting CSV-file from digitalising the location of use-wear traces in QGIS.
The script will:

1. Read in the original CSV-file   
2. Format the data  
3. Calculate the percentages 
4. Plot the data as pie charts 

```{r}
dir_in <- "analysis/raw_data"
dir_out <- "analysis/plots/"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(tools)
library(R.utils)
library(chron)
library(ggplot2)
library(wesanderson)
library(dplyr)
library(doBy)

```


---

# Get names, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.csv$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```

The checksum (MD5 hashes) of the imported file are:  
```{r, echo = FALSE}
info_in

```


---


# Read in original CSV-file
```{r}
imp_data <- read.csv(data_file, header = TRUE, stringsAsFactors = FALSE, na.strings = "*****")
str(imp_data)

```

---


# Percentages
## Percentages of use-wear types for all tool types together 
```{r}
# splits the data into single areas 
sp_all <- split(imp_data, imp_data[["area"]])
pct_all <- vector(mode = "list", length = length(sp_all))


for (i in seq_along(sp_all)) {
  pct_all[[i]] <- sp_all[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_all)[i])
}

PCT_all <- do.call(rbind, pct_all) %>% 
           select(area, use.wear.type, perc)

```


## Percentages of use-wear types for all Keilmesser
```{r}
# defines which part of the data belongs to Keilmesser 
KM <- imp_data[imp_data[["tool.type"]] == "Keilmesser", ]

# splits the data into single areas 
sp_KM <- split(KM, KM[["area"]])
pct_KM <- vector(mode = "list", length = length(sp_KM))


for (i in seq_along(sp_KM)) {
  pct_KM[[i]] <- sp_KM[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_KM)[i])
}

PCT_KM <- do.call(rbind, pct_KM) %>% 
           select(area, use.wear.type, perc)

```


## Percentages of use-wear types for all Keilmesser from Buhlen
```{r}
# defines which part of the data belongs to Keilmesser 
KM <- imp_data[imp_data[["tool.type"]] == "Keilmesser", ]
# selects only the data from Buhlen
KM_Buhlen <- KM [ , ] %>% arrange(site) 
KM_Buhlen <- KM_Buhlen[105:177, ]

# splits the data into single areas 
sp_KM_Buhlen <- split(KM_Buhlen, KM_Buhlen[["area"]])
pct_KM_Buhlen <- vector(mode = "list", length = length(sp_KM_Buhlen))


for (i in seq_along(sp_KM_Buhlen)) {
  pct_KM_Buhlen[[i]] <- sp_KM_Buhlen[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_KM_Buhlen)[i])
}

PCT_KM_Buhlen <- do.call(rbind, pct_KM_Buhlen) %>% 
           select(area, use.wear.type, perc)

```


## Percentages of use-wear types for all Keilmesser from Balve
```{r}
# defines which part of the data belongs to Keilmesser 
KM <- imp_data[imp_data[["tool.type"]] == "Keilmesser", ]
# selects only the data from Balve
KM_Balve <- KM [ , ] %>% arrange(site) 
KM_Balve <- KM_Balve[1:104, ]

# splits the data into single areas 
sp_KM_Balve <- split(KM_Balve, KM_Balve[["area"]])
pct_KM_Balve <- vector(mode = "list", length = length(sp_KM_Balve))


for (i in seq_along(sp_KM_Balve)) {
  pct_KM_Balve[[i]] <- sp_KM_Balve[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_KM_Balve)[i])
}

PCT_KM_Balve <- do.call(rbind, pct_KM_Balve) %>% 
           select(area, use.wear.type, perc)

```


## Percentages of use-wear types for all Keilmesser from Ramioul
```{r}
# defines which part of the data belongs to Keilmesser 
KM <- imp_data[imp_data[["tool.type"]] == "Keilmesser", ]
# selects only the data from Ramioul
KM_Ramioul <- KM [ , ] %>% arrange(site) 
KM_Ramioul <- KM_Ramioul[178:195, ]

# splits the data into single areas 
sp_KM_Ramioul <- split(KM_Ramioul, KM_Ramioul[["area"]])
pct_KM_Ramioul <- vector(mode = "list", length = length(sp_KM_Ramioul))


for (i in seq_along(sp_KM_Ramioul)) {
  pct_KM_Ramioul[[i]] <- sp_KM_Ramioul[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_KM_Ramioul)[i])
}

PCT_KM_Ramioul <- do.call(rbind, pct_KM_Ramioul) %>% 
           select(area, use.wear.type, perc)

```


## Percentages of use-wear types for all Pradnik scraper 
```{r}
# defines which part of the data belongs to Pradnick scraper 
PS <- imp_data[imp_data[["tool.type"]] == "Pradnick scraper", ]

# splits the data into single areas 
sp_PS <- split(PS, PS[["area"]])
pct_PS <- vector(mode = "list", length = length(sp_PS))


for (i in seq_along(sp_PS)) {
  pct_PS[[i]] <- sp_PS[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_PS)[i])
}

PCT_PS <- do.call(rbind, pct_PS) %>% 
           select(area, use.wear.type, perc)

```


## Percentages of use-wear types for all Pradnik spalls
```{r}
# defines which part of the data belongs to Pradnick spall 
LSS <- imp_data[imp_data[["tool.type"]] == "Pradnick spall", ]

# splits the data into single areas 
sp_LSS <- split(LSS, LSS[["area"]])
pct_LSS <- vector(mode = "list", length = length(sp_LSS))


for (i in seq_along(sp_LSS)) {
  pct_LSS[[i]] <- sp_LSS[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_LSS)[i])
}

PCT_LSS <- do.call(rbind, pct_LSS) %>% 
           select(area, use.wear.type, perc)

```


## Percentages of use-wear types for all scraper
```{r}
# defines which part of the data belongs to Pradnick spall 
S <- imp_data[imp_data[["tool.type"]] == "scraper", ]

# splits the data into single areas 
sp_S <- split(S, S[["area"]])
pct_S <- vector(mode = "list", length = length(sp_S))


for (i in seq_along(sp_S)) {
  pct_S[[i]] <- sp_S[[i]] %>% 
              group_by(use.wear.type) %>%
              summarize(perc = round(n() / nrow(.) * 100, digits = 2)) %>% 
              mutate(area = names(sp_S)[i])
}

PCT_S <- do.call(rbind, pct_S) %>% 
           select(area, use.wear.type, perc)

```

# Pie charts
## Colour definitions
```{r}
#05100c black 
#999999 gray 
#52854c green 
#c3d7a4 light green 
#487bb6 blue 
#a6cee3 light blue 
#9a0f0f red
#d16103 orange
#fdbf6f apricot
#ffdb6d yellow
#985633 brown 
#134680 dark blue

custom.col <- data.frame(type = unique(imp_data$use.wear.type), 
                         col = c("#999999", "#52854c", "#c3d7a4", "#487bb6", "#a6cee3",
                                 "#9a0f0f", 
                                 "#d16103", "#ffdb6d", "#985633", "#134680", "#05100c")) 

```


## Plots per tool type 
### Keilmesser 
```{r}
# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp_KM)){
  
  col_i <- custom.col[custom.col$type %in% unique(sp_KM[[i]][["use.wear.type"]]), "col"]
  KM_pie <- ggplot(data = sp_KM[[i]], aes(x = area, fill = use.wear.type )) +
            geom_bar(stat = "count", width = 0.5) + 
            coord_polar("y", start = 0) +
            theme_void() +
            scale_fill_manual(values = col_i) +
            labs(fill = gsub("\\.", " ", "use-wear type"))  
          

 # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_KM_pie_", 
	                   names(sp_KM)[i], ".pdf")
	ggsave(filename = file_out, plot = KM_pie, path = dir_out, device = "pdf")
        
}

```


### Keilmesser from Buhlen
```{r}
# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp_KM_Buhlen)){
  
  col_i <- custom.col[custom.col$type %in% unique(sp_KM_Buhlen[[i]][["use.wear.type"]]), "col"]
  KM_Buhlen_pie <- ggplot(data = sp_KM_Buhlen[[i]], aes(x = area, fill = use.wear.type )) +
                   geom_bar(stat = "count", width = 0.5) + 
                   coord_polar("y", start = 0) +
                   theme_void() +
                   scale_fill_manual(values = col_i) +
                   labs(fill = gsub("\\.", " ", "use-wear type"))  
          

 # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_KM_Buhlen_pie_", 
	                   names(sp_KM_Buhlen)[i], ".pdf")
	ggsave(filename = file_out, plot = KM_Buhlen_pie, path = dir_out, device = "pdf")
        
}

```


### Keilmesser from Balve
```{r}
# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp_KM_Balve)){
  
  col_i <- custom.col[custom.col$type %in% unique(sp_KM_Balve[[i]][["use.wear.type"]]), "col"]
  KM_Balve_pie <- ggplot(data = sp_KM_Balve[[i]], aes(x = area, fill = use.wear.type )) +
                   geom_bar(stat = "count", width = 0.5) + 
                   coord_polar("y", start = 0) +
                   theme_void() +
                   scale_fill_manual(values = col_i) +
                   labs(fill = gsub("\\.", " ", "use-wear type"))  
          

 # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_KM_Balve_pie_", 
	                   names(sp_KM_Balve)[i], ".pdf")
	ggsave(filename = file_out, plot = KM_Balve_pie, path = dir_out, device = "pdf")
        
}

```


### Keilmesser from Ramioul
```{r}
# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp_KM_Ramioul)){
  
  col_i <- custom.col[custom.col$type %in% unique(sp_KM_Ramioul[[i]][["use.wear.type"]]), "col"]
  KM_Ramioul_pie <- ggplot(data = sp_KM_Ramioul[[i]], aes(x = area, 
                                                       fill = use.wear.type )) +
                   geom_bar(stat = "count", width = 0.5) + 
                   coord_polar("y", start = 0) +
                   theme_void() +
                   scale_fill_manual(values = col_i) +
                   labs(fill = gsub("\\.", " ", "use-wear type"))  
          

 # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_KM_Ramioul_pie_", 
	                   names(sp_KM_Ramioul)[i], ".pdf")
	ggsave(filename = file_out, plot = KM_Ramioul_pie, path = dir_out, device = "pdf")
        
}

```


### Pradnick scraper
```{r}
# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp_PS)){
  
  col_i <- custom.col[custom.col$type %in% unique(sp_PS[[i]][["use.wear.type"]]), "col"]
  PS_pie <- ggplot(data = sp_PS[[i]], aes(x = area, fill = use.wear.type )) +
            geom_bar(stat = "count", width = 0.5) + 
            coord_polar("y", start = 0) +
            theme_void() + 
            scale_fill_manual(values = col_i) +
            labs(fill = gsub("\\.", " ", "use-wear type"))  


 # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PS_pie_", 
	                   names(sp_PS)[i], ".pdf")
	ggsave(filename = file_out, plot = PS_pie, path = dir_out, device = "pdf")
        
}

```


### Pradnick spall
```{r}
# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp_LSS)){
  
  col_i <- custom.col[custom.col$type %in% unique(sp_LSS[[i]][["use.wear.type"]]), "col"]
  LSS_pie <- ggplot(data = sp_LSS[[i]], aes(x = area, fill = use.wear.type )) +
             geom_bar(stat = "count", width = 0.5) + 
             coord_polar("y", start = 0) +
             theme_void() + 
             scale_fill_manual(values = col_i) +
             labs(fill = gsub("\\.", " ", "use-wear type"))  
  

 # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_LSS_pie_", 
	                   names(sp_LSS)[i], ".pdf")
	ggsave(filename = file_out, plot = LSS_pie, path = dir_out, device = "pdf")
        
}

```


### Scraper
```{r}
# plots first as a bar plot and then converts into a pie chart  
for (i in seq_along(sp_S)){
  
  col_i <- custom.col[custom.col$type %in% unique(sp_S[[i]][["use.wear.type"]]), "col"]
  S_pie <- ggplot(data = sp_S[[i]], aes(x = area, fill = use.wear.type )) +
             geom_bar(stat = "count", width = 0.5) + 
             coord_polar("y", start = 0) +
             theme_void() + 
             scale_fill_manual(values = col_i) +
             labs(fill = gsub("\\.", " ", "use-wear type"))  
  

 # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_S_pie_", 
	                   names(sp_S)[i], ".pdf")
	ggsave(filename = file_out, plot = S_pie, path = dir_out, device = "pdf")
        
}

```

---

# Save data
## Format name of output file 
```{r}
file_out <- "Use-wear_qualitative_stats"

```

The file will be saved as “~/analysis/summary_stats/.[ext]”.


## Write to XLSX
```{r}
write.xlsx(list(all = PCT_all, KM = PCT_KM, PCT_PS = PS, PCT_LSS = LSS, PCT_S = S), 
                file = paste0(dir_out, file_out, ".xlsx"))

```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
